---
title: "Lab09"
author: "Elly Jang, A18582705, rsjang@ucsd.edu"
format: pdf
editor: visual
---

**IMPORTING CANDY DATA**

```{r}
candy_file <- "https://raw.githubusercontent.com/fivethirtyeight/data/master/candy-power-ranking/candy-data.csv"

candy = read.csv(candy_file, row.names=1)
head(candy)
```

**Q1. How many different candy types are in this dataset?**

> **85**

**Q2. How many fruity candy types are in the dataset?**

> **38**

```{r}
#Q1. Number of candy types
dim(candy)

#Q2. Number of fruity candy
sum(candy$fruity == 1)
```

**Q3. What is your favorite candy (other than Twix) in the dataset and what is it’s winpercent value?**

> **Milky Way, winpercent = 73.09956**

**Q4. What is the winpercent value for “Kit Kat”?**

> **76.7686**

**Q5. What is the winpercent value for “Tootsie Roll Snack Bars”?**

> **49.6535**

```{r}
library(dplyr)

#Q3. winpercent of favorite candy, "Milky Way"
candy |> 
  filter(row.names(candy)=="Milky Way") |> 
  select(winpercent)

#Q4. winpercent of "Kit Kat"
candy |> 
  filter(row.names(candy)=="Kit Kat") |> 
  select(winpercent)

#Q5. Winpercent of "Tootsie Roll Snack Bars"
candy |> 
  filter(row.names(candy)=="Tootsie Roll Snack Bars") |> 
  select(winpercent)
```

**Q6. Is there any variable/column that looks to be on a different scale to the majority of the other columns in the dataset?**

> **Yes. Most variables are on the 0-1 scale, but "winpercent" stands out because it is a continuous numeric variable.**

**Q7. What do you think a zero and one represent for the candy\$chocolate column?**

> **In the chocolate column, 0 indicates the candy does not contain chocolate, and the 1 indicates that the candy does contain chocolate.**

```{r}
library("skimr")
skim(candy)

#Q6 code
summary(candy)

#Q7 code
table(candy$chocolate)
```

**EXPLORATORY ANALYSIS**

**Q8. Plot a histogram of winpercent values using both base R an ggplot2.**

```{r}
#Q8. Histogram of winpercent...
##Method 1: Using Base R
hist(candy$winpercent, main = "Distribution of Winpercent",
     xlab = "Winpercent", col = "blue", breaks = 23)

##Method 2: Using ggplot2
library(ggplot2)
ggplot(candy, aes(x=winpercent)) +
  geom_histogram(bins = 23, fill = "blue", color = "black") +
  labs(title = "Distribution of Winpercent",
       x = "Winpercent", y = "Count")
```

**Q9. Is the distribution of winpercent values symmetrical?**

> **No, it seems kind of skewed.**

**Q10. Is the center of the distribution above or below 50%?**

> **The center of distribution is above 50%, as the mean (50.3%) is above 50%.**

```{r}
#Q10 Center of distribution above or below?
mean(candy$winpercent)
median(candy$winpercent)
```

**Q11. On average is chocolate candy higher or lower ranked than fruit candy?**

> **On average, chocolate candies are ranked higher than fruit candies.**

```{r}
#Q11 Code
mean(candy$winpercent[as.logical(candy$chocolate)])
mean(candy$winpercent[as.logical(candy$fruity)])

```

**Q12. Is this difference statistically significant?**

> **Since the p-value is 2.871e-08, which is \<0.05, the difference is statistically significant.**

```{r}
#Q12 t.test
t.test(candy$winpercent[as.logical(candy$chocolate)],
       candy$winpercent[as.logical(candy$fruity)])
```

**OVERALL CANDY RANKINGS**

**Q13. What are the five least liked candy types in this set?**

> **Nik L Nip, Boston Baked Beans, Chiclets, Super Bubble, and Jawbusters.**

```{r}
#Q13 Code using base R
head(candy[order(candy$winpercent),], n = 5)

#Q13 Code using dplyr
candy |> arrange(winpercent) |> head(5)
```

**Q14. What are the top 5 all time favorite candy types out of this set?**

> **Reese's Peanut Butter cup, Reese's Miniatures, Twix, Kit Kat, and Snickers.**

```{r}
#Q14 Code using base R
head(candy[order(candy$winpercent, decreasing = TRUE),], n = 5)

#Q14 Code using dplyr
candy |> arrange(desc(winpercent)) |> head(5)
```

**Q15. Make a first barplot of candy ranking based on winpercent values.**

```{r}
#Q15 barplot
ggplot(candy, aes(x = winpercent, y = rownames(candy))) +
  geom_col()
```

**Q16. This is quite ugly, use the reorder() function to get the bars sorted by winpercent.**

```{r}
#Q16 recorder bars to sort by winpercent
ggplot(candy, aes(winpercent, reorder(rownames(candy), winpercent))) +
  geom_col()
```

```{r}
#Add color
my_cols=rep("black", nrow(candy))
my_cols[as.logical(candy$chocolate)] = "chocolate"
my_cols[as.logical(candy$bar)] = "brown"
my_cols[as.logical(candy$fruity)] = "pink"

ggplot(candy) + 
  aes(winpercent, reorder(rownames(candy),winpercent)) +
  geom_col(fill=my_cols) 
```

**Q17. What is the worst ranked chocolate candy?**

> **The worst ranked chocolate candy is "Sixlets".**

**Q18. What is the best ranked fruity candy?**

> **The best ranked fruity candy is "Starburst".**

**TAKING A LOOK AT PRICEPERCENT**

```{r}
library(ggrepel)

# How about a plot of win vs price
ggplot(candy) +
  aes(winpercent, pricepercent, label=rownames(candy)) +
  geom_point(col=my_cols) + 
  geom_text_repel(col=my_cols, size=3.3, max.overlaps = 5)
```

**Q19. Which candy type is the highest ranked in terms of winpercent for the least money - i.e. offers the most bang for your buck?**

> **Reese's Peanut Butter Cup**

```{r}
#Q19 Ranking in terms of winpercent for the least money
candy$value_score <- candy$winpercent - candy$pricepercent
candy |> arrange(desc(value_score)) |> head(1)
```

**Q20. What are the top 5 most expensive candy types in the dataset and of these which is the least popular?**

> **Top 5 most expensive candy types: Nik L Nip, Nestle Smarties, Ring pop, Hershey's Krackel, and Hershey's Milk Chocolate. The least popular among these is Nik L Nip.**

```{r}
#Q20 Price of Candy
ord <- order(candy$pricepercent, decreasing = TRUE)
head(candy[ord,c(11,12)], n=5)
```

**EXPLORING THE CORRELATION STRUCTURE**

```{r}
library(corrplot)

candy_BK <- candy
candy <- candy %>% dplyr::select(-value_score)
cij <- cor(candy)
corrplot(cij)
```

**Q22. Examining this plot, what two variables are anti-correlated (i.e. have minus values)?**

> **There is a negative correlation between "fruity" and "chocolate".**

**Q23. Similarly, what two variables are most positively correlated?**

> **There is a positive correlation between "chocolate" and "bar".**

**PRINCIPAL COMPONENT ANALYSIS**

```{r}
pca <- prcomp(candy, scale = TRUE)
summary(pca)

plot(pca$x[,1], pca$x[,2])

plot(pca$x[,1:2], col=my_cols, pch=16)
```

```{r}
#Nicer plot with ggplot2
##Make a new data-frame with our PCA results and candy data
my_data <- cbind(candy, pca$x[,1:3])

p <- ggplot(my_data) + 
        aes(x=PC1, y=PC2, 
            size=winpercent/100,  
            text=rownames(my_data),
            label=rownames(my_data)) +
        geom_point(col = my_cols)

p
```

```{r}
#Nicer plot with ggrepel
library(ggrepel)

p + geom_text_repel(size=3.3, col=my_cols, max.overlaps = 7)  + 
  theme(legend.position = "none") +
  labs(title="Halloween Candy PCA Space",
       subtitle="Colored by type: chocolate bar (dark brown),
       chocolate other (light brown), fruity (red), other
       (black)",
       caption="Data from 538")
```

**Q24. Complete the code to generate the loadings plot above. What original variables are picked up strongly by PC1 in the positive direction? Do these make sense to you? Where did you see this relationship highlighted previously?**

> **Original variables that are picked up strongly by PC1 in the positive direction: fruity and pluribus (and hard). Yes, this does make sense because it represents a contrast between the two candy types: chocolate and fruity. We've seen this relationship in previous questions: in Q22, the "chocolate" and "fruity" variables were negatively correlated, and again in Q11/12, chocolate candies were ranked higher than fruity candies, which suggests the two types perform differently from each other.**

```{r}
library(plotly)
ggplotly(p)

#Completed Q24 code
loadings <- as.data.frame(pca$rotation)

ggplot(loadings) +
  aes(x = PC1, y = reorder(rownames(loadings), PC1)) +
  geom_col()
```

**SUMMARY**

**Q25. Based on your exploratory analysis, correlation findings, and PCA results, what combination of characteristics appears to make a “winning” candy? How do these different analyses (visualization, correlation, PCA) support or complement each other in reaching this conclusion?**

> **A "winning" candy appears to contain chocolate, often in bar form and paired with additions, such as nuts, caramel, or nougat, rather than a simple fruity candy. The visual analysis showed that chocolate candies consistently ranked higher compared to fruity candies. The correlation analysis supports this finding that chocolate and fruity are strongly negatively correlated. The PCA integrates these patterns, with PC1 separating chocolate-based candies and fruity-based candies, which indicates that this is the main source of variation.**
