---
title: "Lab10"
author: "Elly Jang, A18582705, rsjang@ucsd.edu"
format: pdf
editor: visual
---

**PDB STATISTICS**

**Q1: What percentage of structures in the PDB are solved by X-Ray and Electron Microscopy.**

> **A1. X-Ray: about 80.95077 and EM: about 12.83843.**

```{r}
pdb <- read.csv("pdb_stats.csv")

#Column names
str(pdb)

total <- sum(pdb$Total)
xray <- sum(pdb$X.ray)
em <- sum(pdb$EM)

xray1 <- xray / total * 100
em1 <- em / total * 100

xray1
em1

```

**Q2. What proportion of structures in the PDB are protein?**

> **A2. About 97.91%**

```{r}
protein <- sum(pdb$Total[1:3])
total <- sum(pdb$Total)

protein_prop <- protein / total
protein_prop
```

**Q3: Type HIV in the PDB website search box on the home page and determine how many HIV-1 protease structures are in the current PDB?**

> **A3. About 1,147.**

**Q4. Water molecules normally have 3 atoms. Why do we see just one atom per water molecule in this structure?**

> **A4. It could be due to a x-ray diffraction limitation. Or due to size of the atoms: hydrogens are small and thus not as visible, but the oxygen atom is larger and thus easier to detect.**

**Q5. There is a critical "conserved water molecule in the binding site. Can you identify this water molecule? What residue number does this water molecule have?**

> **A5. Yes. It is between the two ASP 25 residues, where it helps stabilize ligand binding. Residue number is HOH 308.**

**Q6. Generate and save a figure clearly showing the two distinct chains of HIV-protease along with the ligand. You might also consider showing the catalytic residues ASP 25 in each chain and the critical water (we recommend “Ball & Stick” for these side-chains). Add this figure to your Quarto document.**

> **A6. *Done below***

![1HSG ASP 25](1HSG(3).png)

**INTRODUCTION TO BIO3D IN R**

```{r}
library(bio3d)
pdb <- read.pdb("1hsg")
pdb
```

**Q7. How many amino acid residues are there in this pdb object?**

> **198**

**Q8. Name one of the two non-protein residues.**

> **MK1**

**Q9. How many protein chains are in this structure?**

> **2**

```{R}
attributes(pdb)
```

```{r}
head(pdb$atom)
```

```{r}
library(bio3dview)
library(NGLVieweR)

view.pdb(pdb) |>
  setSpin()
```

```{r}
#Select the important ASP 25 reside
sele <- atom.select(pdb, resno=25)

#and highlight them in spacefill representation
view.pdb(pdb, cols=c("navy", "teal"),
         highlight = sele,
         highlight.style = "spacefill") |>
  setRock()
```

**PREDICTING FUNCTIONAL MOTIONS OF A SINGLE STRUCTURE**

```{r}
adk <- read.pdb("6s36")
adk
```

```{r}
# Perform flexiblity prediction
m <- nma(adk)
plot(m)
```

```{r}
mktrj(m, file = "adk_m7.pdb")

view.nma(m, pdb=adk)
```

**COMPARATIVE STRUCTURE ANALYSIS OF ADENYLATE KINASE**

**Q10. Which of the packages above is found only on BioConductor and not CRAN?**

> **A10. msa**

**Q11. Which of the above packages is not found on BioConductor or CRAN?**

> **A11. bio3dview**

**Q12. True or False? Functions from the pak package can be used to install packages from GitHub and BitBucket.**

> **A12. TRUE**

**SEARCH AND RETRIEVE ADK STRUCTURES**

```{r}
library(bio3d)
aa <- get.seq("1ake_A")
aa
```

**Q13. How many amino acids are in this sequence, i.e. how long is this sequence?**

> **A13. 214**

```{r}
#BLAST search
b <- blast.pdb(aa)
hits <- plot(b)

#Top hits
head(hits$pdb.id)

#"Required"
hits <- NULL
hits$pdb.id <- c('1AKE_A','6S36_A','6RZE_A','3HPR_A','1E4V_A','5EJE_A','1E4Y_A','3X2S_A','6HAP_A','6HAM_A','4K46_A','3GMT_A','4PZL_A')
```

**ANNOTATE COLLECTED PDB STRUCTURES**

```{r}
#Download related PDB files
files <- get.pdb(hits$pdb.id, path="pdbs", split = TRUE, gzip = TRUE)

#Align related PDBs
pdbs <- pdbaln(files, fit = TRUE, exefile = "msa")
```

```{r}
#Vector containing PDB database codes
ids <- basename.pdb(pdbs$id)

anno <- pdb.annotate(ids)
unique(anno$source)

anno
```

**PRINCIPAL COMPONENT ANALYSIS**

```{r}
# Perform PCA
pc.xray <- pca(pdbs)
plot(pc.xray)
```

```{r}
#Calculate RMSD
rd <- rmsd(pdbs)

#Structure-based clustering
hc.rd <- hclust(dist(rd))
grps.rd <- cutree(hc.rd, k=3)

plot(pc.xray, 1:2, col="grey50", bg=grps.rd, pch=21, cex=1)
```

**PCA VISUALIZATION**

```{r}
#Visualize first principal component
pcl <- mktrj(pc.xray, pc=1, file = "pc_1.pdb")
```

```{r}
#PCA results with ggplot
library(ggplot2)
library(ggrepel)

df <- data.frame(PC1=pc.xray$z[,1],
                 PC2=pc.xray$z[,2],
                 col=as.factor(grps.rd),
                 ids=ids)

p <- ggplot(df) +
  aes(PC1, PC2, col=col, label=ids) +
  geom_point(size=2) +
  geom_text_repel(max.overlaps = 20) +
  theme(legend.position = "none")

p
```
